Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-01 fib2.tex
This version compiles for me.
I changed to pdf figures since there is a problem with eps2pdf package, and I put the tikz figures back in, but smaller.
Tried to make the fusion and braid diagrams a little more compact.
Copied and pasted the tikz code from Jiannis's paper in case we want to use that instead.
Removed Jiannis diagrams and changed "bond cluster" to "connected component"

%\usepackage{tikz}
%\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit}

@@

The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
	\begin{align}
		\begin{tikzpicture}[baseline=0.5cm,scale=0.33]		
			\draw(0,0)--(0,1);
			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\node at (0,0) [anchor=north] {$c$};
			\node at (-1,3) [anchor=south] {$b$};
			\node at (1,3) [anchor=south] {$a$};
		\end{tikzpicture}
		&\;\;=\;\;R^{ab}_c\cdot\;
		\begin{tikzpicture}[baseline=0.5cm,scale=0.33]		
			\draw(0,0)--(0,1);
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (0,0) [anchor=north] {$c$};
			\node at (-1,3) [anchor=south] {$b$};
			\node at (1,3) [anchor=south] {$a$};
		\end{tikzpicture}
	\end{align}
where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, denoted as
	\begin{align}
		\begin{tikzpicture}[baseline=0.6cm,scale=0.4]
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(-1,2);
			\draw(2,3)--(0,1);
			\node at (-0.45,1.6) [anchor=north east] {$i$};
			\node at (0,0) [anchor=north] {$d$};
			\node at (2,3) [anchor=south] {$c$};
			\node at (0,3) [anchor=south] {$b$};
			\node at (-2,3) [anchor=south] {$a$};
		\end{tikzpicture}
		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
		\begin{tikzpicture}[baseline=0.6cm,scale=0.4]		
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(1,2);
			\draw(2,3)--(0,1);
			\node at (0.42,1.63) [anchor=north west] {$j$};
			\node at (0,0) [anchor=north] {$d$};
			\node at (2,3) [anchor=south] {$c$};
			\node at (0,3) [anchor=south] {$b$};
			\node at (-2,3) [anchor=south] {$a$};
		\end{tikzpicture}
	\end{align}

\dude{these equations are too big... i was hoping for something tight, like
in http://arxiv.org/abs/1504.02868
any ideas? i might just generate a bunch of little pdf's...
}

@@

Our error-correction algorithm is
based on a heirarchical clustering algorithm~\cite{Hastie2009},
and has some 
similarities to the hard-decision renormalization group (RG) decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and 
%fusing the anyons within each cluster. 
measureing the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales.
See FIG. \ref{f:decode}.
Note that unlike the abelian case, the decoder
cannot know the charge of each cluster based only
on the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
The simulation proceeds in this way as a dialogue between decoder
and system, terminating when the lattice 
is free of anyons (i.e.~we have returned to the codespace) 
or the decoder performs a homologically nontrivial operation, thereby altering the codespace.
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-decode.eps}
\caption{The decoder works by maintaining a set of disjoint clusters as rooted trees.
(a) Trees are formed from neighboring sites that contain a charge.
(b) For each successive round, we grow all trees, 
and overlapping trees are joined to form
a single tree. Arrows denote movement of charge to the root of the tree.
The root of each tree is chosen arbitrarily.
}
\label{f:decode}
\end{center}
\end{figure}

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate clusters~\footnote{
We use the word cluster to refer both to groupings 
of nearby anyons as formed by the decoding algorithm and 
to connected components of a graph under percolation\dude{confusing}, since both 
of these usages are standard in the literature. The intended 
meaning will remain clear from context.} of average size $O(\log(n))$ and variance 
$O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected clusters correspond to sets of anyons that could not have interacted at any 
point in their history. 
Although normally anyons can interact by braiding without ever coming 
into contact with one another \dude{um um um}, in this case charge conservation 
implies that each distinct cluster must have trivial total charge, and so all such interactions are trivial.
This allows us to simulate the braiding processes within each cluster 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over bond clusters, and we can compute each factor separately. Since each 
cluster has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.eps}
\caption{Four equivalent representations of a state,
along with two syndrome observables marked with solid and dashed lines:
(a) two pairs are joined in (b).
Deform the curve straight to get (c),
then effect some
braid operations on the state so that (d) the two
observables are now represented on the curve. 
%\cggb{I'm not totally clear on what I'm looking at here. Is this the diagram where we want to explain the whole simulation algorithm? Is the idea that we start with a), which is given by the set of noise processes that have happened, then we connect stuff up arbitrarily into b) to make some fusion tree. Then c) is what's happening after linearizing the picture? And d) is c) after some R and F moves so you go into the standard basis for the fusion tree? Are the red lines between c) and d) just denoting leaves with the same identity or do they have some more precise meaning wrt the braid moves that relate c) and d)?}
}
\label{f:syndrome}
\end{center}
\end{figure}

>>>>

\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit,shapes.misc}

@@

The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
	\begin{align}
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;R^{ab}_c\cdot\;
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}
where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, denoted as
	\begin{align}
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(-1,2);
			\draw(2,3)--(0,1);
			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(1,2);
			\draw(2,3)--(0,1);
			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009},
and has some 
similarities to the hard-decision renormalization group (RG) decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and 
%fusing the anyons within each cluster. 
measuring the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales.
See FIG. \ref{f:decode}.
Note that unlike the abelian case, the decoder
cannot know the charge of each cluster based only
on the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
The simulation proceeds in this way as a dialogue between decoder
and system, terminating when the lattice 
is free of anyons (i.e.~we have returned to the codespace) 
or the decoder performs a homologically nontrivial operation, thereby altering the codespace.
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-decode.pdf}
\caption{The decoder works by maintaining a set of disjoint clusters as rooted trees.
(a) Trees are formed from neighboring sites that contain a charge.
(b) For each successive round, we grow all trees, 
and overlapping trees are joined to form
a single tree. Arrows denote movement of charge to the root of the tree.
The root of each tree is chosen arbitrarily.
}
\label{f:decode}
\end{center}
\end{figure}

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
%~\footnote{We use the word cluster to refer both to groupings of nearby anyons as formed by the decoding algorithm and to connected components of a graph under percolation\dude{confusing}, since both of these usages are standard in the literature. The intended meaning will remain clear from context.}
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
Although normally anyons can interact by braiding without ever coming 
into contact with one another \dude{um um um}, in this case charge conservation 
implies that each distinct component must have trivial total charge, and so all such interactions are trivial.
This allows us to simulate the braiding processes within each component 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components, and we can compute each factor separately. Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}
\caption{Four equivalent representations of a state,
along with two syndrome observables marked with solid and dashed lines:
(a) two pairs are joined in (b).
Deform the curve straight to get (c),
then effect some
braid operations on the state so that (d) the two
observables are now represented on the curve. 
%\cggb{I'm not totally clear on what I'm looking at here. Is this the diagram where we want to explain the whole simulation algorithm? Is the idea that we start with a), which is given by the set of noise processes that have happened, then we connect stuff up arbitrarily into b) to make some fusion tree. Then c) is what's happening after linearizing the picture? And d) is c) after some R and F moves so you go into the standard basis for the fusion tree? Are the red lines between c) and d) just denoting leaves with the same identity or do they have some more precise meaning wrt the braid moves that relate c) and d)?}
}
\label{f:syndrome}
\end{center}
\end{figure}
